---
title: "Project 1"
author: "Stephanie Bruce"
output:
  html_document:
    keep_md: yes
  pdf_document: default
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Reproducible Research - Course Project 1 

This R Markdown document include my code for Project 1.

```{r read-data, echo=TRUE}
library(readr)
data <- read_csv("/home/rstudio/Reproducible_Research/week2/activity.csv")
```

## Part 1. What is mean total number of steps taken per day 
```{r part1a, echo=TRUE}
total_steps <- aggregate(x=data$steps, FUN=sum, by=list(Group.date = data$date), na.rm=TRUE)
```
The total number of steps taken per day is:  
```{r part1a1, echo=TRUE}
print(total_steps)
```

```{r part1b, echo=TRUE}
plot(total_steps$Group.date, total_steps$x, type = "h") 
title(main = "Histogram of steps by Day")
```

The mean and median number of steps per day are: 
```{r part1c, echo=TRUE}
mean_steps <- mean(total_steps$x)
median_steps <- median(total_steps$x)
print(mean_steps)
print(median_steps)
```

## Part 2. What is the average daily activity pattern 
```{r part2, echo=TRUE}
interval_totals <- aggregate(x=data$steps, FUN=sum, by=list(group.date = data$interval), na.rm=TRUE)
mean_interval <-aggregate(x=data$steps, FUN=mean, by=list(group.date = data$interval), na.rm=TRUE)
```

Time series plot by interval
```{r part2b, echo=TRUE}
plot(mean_interval$group.date, mean_interval$x, type = "l")
title(main = "Mean Steps by 5-minute interval throughout day")
```
The average maximum number of steps is at interval 825 and the max steps is:
```{r part2c, echo=TRUE}
print(max(mean_interval$x))
```

The daily average number of steps is:
```{r part2extra, echo=TRUE}
daily_mean <- sum(mean_interval$x)
```

## Part 3. Imputing missing values 
Calculate and report the total number of missing values in the dataset:
```{r part3, echo=TRUE}
count_nas <- aggregate(x=data$steps, function(x) {sum(is.na(x))}, by=list(Group.date = data$date))
print(sum(count_nas$x))
```
Fill in all the missing values in the dataset (I split the daily average into each interval):
```{r part3a1, echo=TRUE}
data2 <- data
data2$impute <- data$steps
data2$impute[is.na(data2$impute)] <- daily_mean/length(mean_interval$x)
```

Create a new dataset that is equal to the original dataset (new dataset = data2):
```{r part3b, echo=TRUE}
total_impute <- aggregate(x=data2$impute, FUN=sum, by=list(Group.date = data$date))
```
Histogram:
```{r part3c, echo=TRUE}
plot(total_impute$Group.date, total_impute$x, type = "h") 
title(main = "Histogram of steps by Day - Imputed")
```
Mean and Median steps with imputed data:
```{r part3d, echo=TRUE}
mean_steps_imp <- mean(total_impute$x)
median_steps_imp <- median(total_impute$x)
print(mean_steps_imp)
print(median_steps_imp)
```

## Part 4. Are there differences in activity patterns between weekdays and weekends? 
Create a new factor
```{r part4, echo=TRUE}
dataDoW <- weekdays(data2$date)
weekdays <- c('Monday', 'Tuesday', 'Wednesday', 'Thursday','Friday')
data2$WDay <- factor((weekdays(data$date) %in% weekdays), levels = c(TRUE, FALSE), labels= c('weekday', 'weekend'))
```
Finishing defining weekday vs weekend
```{r part4a, echo=TRUE}
data2$weekday = ifelse(data2$WDay == "weekday", 1,0)
data2$weekend = ifelse(data2$WDay == "weekend", 1,0)
```
Adding information to dataset:
```{r part4a2, echo=TRUE}
data2$wkdy_steps = data2$impute*data2$weekday
data2$wked_steps = data2$impute*data2$weekend
```
Combining information
```{r part4b, echo=TRUE}
interval_wkday <- aggregate(x=data2$wkdy_steps, FUN=sum, by=list(group.date = data2$interval))
interval_wkend <-aggregate(x=data2$wked_steps, FUN=sum, by=list(group.date = data2$interval))
```
Remove 0 out of average from weekend/weekday:
```{r part4b2, echo=TRUE}
interval_wkd <- interval_wkday$x/(sum(data2$weekday)/288)
interval_wke <- interval_wkend$x/(sum(data2$weekend)/288)
```
Panel Plot
```{r part4bc, echo=TRUE}
par(mfrow=c(2,1))
plot(interval_wkday$group.date, interval_wkd, type = "l", ylim = c(0,200), xlab = "interval", ylab = "Weekday")
title("Mean Steps by 5-min interval Weekday vs Weekend")
plot(interval_wkday$group.date, interval_wke, type = "l", ylim=c(0,200), xlab="interval", ylab = "Weekend")
```       
```{r}
```
The End
---